:::{.callout-note}
:::

# Data description

Our dataset contains text data about reviews from the Internet Movie Database (IMDB). Specifically, it consists of 5000 movie reviews (selected for sentiment analysis). No single movie has more than 30 reviews and the sentiment of the reviews is given as a binary variable, i.e. an IMDB rating < 5 results in a sentiment score of 0, and a rating >=7 results in a sentiment score of 1.

```{r}
#| label: R packages
#| echo: false
#| warning: false
#| message: false

library(text2vec)
library(tidyverse)
library(tidytext)  # for text mining
library(tm)        # needed for wordcloud
library(wordcloud) # to create pretty word clouds
library(SnowballC) # for Porter's stemming algorithm
library(ggplot2)
library(cluster)   # for the silhouette score
library(clusterSim) # for the DB-index
library(mclust) # for the adjustedRand index
library(aricode) # for the adjustedRand index
library(magrittr)
library(NLP)         
library(RColorBrewer)
library(umap)
library(mclust)
library(lda)
library(topicmodels)

```

```{r}
#| label: data loading

data("movie_review")

```

```{r}
#| label: eda visualization
#| warning: false

# Plotting the word-cloud plot with the 50 most frequent words with a minimum frequency of 5
word_cloud <- wordcloud(
  words = movie_review$review,
  min.freq = 5,
  max.words = 50,
  random.order = FALSE,
  colors = brewer.pal(8, "Dark2")
)

print(word_cloud)

positive_negative_barplot <- ggplot(movie_review, aes(x = factor(sentiment))) +
  geom_bar(fill = c("red", "green")) +
  labs(title = "Sentiment Distribution of Movie Reviews", x = "Sentiment", y = "Count") +
  scale_x_discrete(labels = c("0" = "Negative", "1" = "Positive")) +
  theme_minimal()
print(positive_negative_barplot)

```
From the word cloud graph, we can see that the most common words with font size correspond to word frequency.
As for the barplot of positive-negative reviews, we can conclude that the number of positive reviews is almost the same as the number of negative reviews in our movie_review dataset.

# Text pre-processing

In our case, after loading the dataset, we used a series of text pre-processing steps to clean and standardize the data in preparation for analysis. First, we created a corpus object, which allows us to apply text pre-processing functions to the entire set of movie reviews. We then made the text uniform by converting it all to lowercase, ensuring that all words are treated the same. Next, we removed punctuation to eliminate all punctuation marks from the text and reduce noise. We also removed stopwords—both standard English stopwords like "and" and "the" that contribute little to the context, and custom stopwords specific to our dataset, such as "film," "movie," and "watch," which are too common or irrelevant for our analysis. Finally, we removed any extra whitespace to create cleaner and more readable text. These pre-processing steps help prepare the data for more effective analysis by focusing on the most meaningful terms in the reviews.


```{r}
#| label: Pre-processing steps

# Create a Corpus from the review text
corpus <- Corpus(VectorSource(movie_review$review))

# Custom stop words
custom_stopwords <- c("film", "movie", "watch")
all_stopwords <- c(stopwords("en"), custom_stopwords)

# Preprocess the Corpus
corpus <- tm_map(corpus, content_transformer(tolower)) # Lowercasing
corpus <- tm_map(corpus, removePunctuation) # Punctuation removal
corpus <- tm_map(corpus, removeNumbers) # Number removal
corpus <- tm_map(corpus, removeWords, all_stopwords) # Stopword removal
corpus <- tm_map(corpus, stripWhitespace) # Remove extra whitespace

```

# Text representation

A Document-Term Matrix (DTM) is created to organize the text data by showing how often each word appears in each document, making it easier to analyze. Only words with at least three characters are included, filtering out very short, less meaningful words. Additionally, words that appear in fewer than 5% of the documents are removed, focusing on more common and relevant terms. The DTM is then converted into a matrix and standardized so that each word has a similar impact in the analysis. This structured matrix highlights the most frequent words, making it useful for uncovering patterns or insights in the text data.

```{r}
#| label: DTM
# Create a Document-Term Matrix (DTM) with frequent terms only
dtm <- DocumentTermMatrix(corpus, control = list(word_length = c(3, Inf))) # Min length 3
dtm <- removeSparseTerms(dtm, 0.95)  # Retain terms present in at least 5% of documents
dtm_matrix <- as.matrix(dtm)

# Scale the DTM matrix to standardize data
dtm_matrix <- scale(dtm_matrix)

inspect(dtm)


```



# Text clustering

In order to analyze the movie review dataset, we used several text representation methods, such as k-Means, k-Medians, GMM, Hierarchical Clustering and Latent Dirichlet Allocation (LDA). The evaluation of clustering was performed using silhouette scores, which measure how similar each document is to its assigned cluster compared to other clusters as well as the Davies-Bouldin Index for internal validation. 

First, we apply K-means algorithm to our dataset, which randomly assigns data points to K clusters and then it calculates the centroids for each cluster. Finally, it assigns each data point to the cluster belonging to its nearest centroid. If the assignments change, it recalculates the centroids based on the newly formed clusters.

```{r}
#| label: K-Means Method
# Create a Document-Term Matrix with only the most frequent terms
dtm <- DocumentTermMatrix(corpus, control = list(word_length = c(3, Inf)))  # Minimum word length of                                                                               3

# Apply tf-idf transformation
dtm_tfidf <- weightTfIdf(dtm)
inspect(dtm_tfidf)

# Remove the sparse terms
dtm_tfidf <- removeSparseTerms(dtm_tfidf, 0.95) # Keep terms that appear in at least 5% of the                                                         documents
dtm_tfidf_m <- as.matrix(dtm_tfidf)


set.seed(40) # for reproducibility
### K-Means Clustering with K=5 ###
###################################
K5 <- 5 # number of clusters

# Sample a subset of the data (e.g., 1000 documents)
#sample_indices <- sample(1:nrow(review_DTM_scaled), 1000)
#review_DTM_sample <- review_DTM[sample_indices,]

# Run K-means on the sample
kmeans_result_5 <- kmeans(dtm_tfidf_m, centers = K5, nstart = 25)

# Add the cluster assignment to the original data
movie_review$Cluster5 <- kmeans_result_5$cluster

# Bar plot showing the number of reviews in each cluster
barplot_kmeans5 <- ggplot(movie_review, aes(x = factor(Cluster5), fill = factor(sentiment))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("red", "green"), 
                    labels = c("Negative", "Positive")) +  # Adjust the labels based on your binary                                                               encoding
  labs(title = "Distribution of Sentiments Across the 5 Clusters",
       x = "Cluster", 
       y = "Number of Reviews", 
       fill = "Sentiment") +
  theme_minimal()

print(barplot_kmeans5)

### K-Means Clustering with K=10 ###
###################################
K10 <- 10 # number of clusters

# Run K-means on the sample
kmeans_result_10 <- kmeans(dtm_tfidf_m, centers = K10, nstart = 25)

# Add the cluster assignment to the original data
movie_review$Cluster10 <- kmeans_result_10$cluster

# Bar plot showing the number of reviews in each cluster
barplot_kmeans10 <- ggplot(movie_review, aes(x = factor(Cluster10), fill = factor(sentiment))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("red", "green"), 
                    labels = c("Negative", "Positive")) +  # Adjust the labels based on your binary                                                               encoding
  labs(title = "Distribution of Sentiments Across the 10 Clusters",
       x = "Cluster", 
       y = "Number of Reviews", 
       fill = "Sentiment") +
  theme_minimal()

print(barplot_kmeans10)

##### Silhouette Score (Internal validation) ########
#####################################################
# Calculate Silhouette score for K=5
silhouette_score_5 <- silhouette(kmeans_result_5$cluster, dist(dtm_tfidf_m))

# View average silhouette width (for K=5)
avg_silhouette_5 <- mean(silhouette_score_5[, 3])
print(paste("Average silhouette score (K=5): ", round(avg_silhouette_5, 2)))

# Calculate Silhouette score for K=10
silhouette_score_10 <- silhouette(kmeans_result_10$cluster, dist(dtm_tfidf_m))

# View average silhouette width (for K=10)
avg_silhouette_10 <- mean(silhouette_score_10[, 3])
print(paste("Average silhouette score (K=10): ", round(avg_silhouette_10, 2)))


#### Davies-Bouldin (DB) index (Internal validation) #####
#########################################
# Calculate Davies-Bouldin Index for K=5
db_index_result_5 <- index.DB(dtm_tfidf_m, kmeans_result_5$cluster)

# View DB index result
db_index_5 <- db_index_result_5$DB
print(paste("Davies-Bouldin Index (K=5): ", round(db_index_5, 2)))

# Calculate Davies-Bouldin Index for K=10
db_index_result_10 <- index.DB(dtm_tfidf_m, kmeans_result_10$cluster)

# View DB index result
db_index_10 <- db_index_result_10$DB
print(paste("Davies-Bouldin Index (K=10): ", round(db_index_10, 2)))

```


Next, we apply the K-Medians algorithm, which initially assigns data points to K clusters randomly. It then calculates the medians for each cluster instead of centroids. Each data point is assigned to the cluster with the closest median, measured using Manhattan distance.

```{r}
#| label: K-Median Method

set.seed(40) # for reproducibility
# Define clusters
k1 <- 5
k2 <- 10
 
# K-Medians clustering with 5 and 10 clusters
k_median_result1 <- pam(dtm_matrix, k1, metric = "manhattan")
k_median_result2 <- pam(dtm_matrix, k2, metric = "manhattan")
 
# Add cluster assignments to movie_review data
movie_review$cluster1 <- k_median_result1$clustering
movie_review$cluster2 <- k_median_result2$clustering
 
# Plot distribution of sentiments for 5 clusters
ggplot(movie_review, aes(x = factor(cluster1), fill = factor(sentiment))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("red", "green"), 
                    labels = c("Negative", "Positive")) +
  labs(title = "Distribution of Sentiments Across Clusters (k = 5)",
       x = "Cluster", 
       y = "Number of Reviews", 
       fill = "Sentiment") +
  theme_minimal()
 
# Plot distribution of sentiments for 10 clusters
ggplot(movie_review, aes(x = factor(cluster2), fill = factor(sentiment))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("red", "green"), 
                    labels = c("Negative", "Positive")) +
  labs(title = "Distribution of Sentiments Across Clusters (k = 10)",
       x = "Cluster", 
       y = "Number of Reviews", 
       fill = "Sentiment") +
  theme_minimal()
 
 
# Calculate distance matrix based on the DTM matrix
dist_matrix <- dist(dtm_matrix, method = "manhattan")  # Use "manhattan" if it matches your clustering metric
 
# Silhouette analysis for k = 5 clusters
silhouette_result_k1 <- silhouette(k_median_result1$clustering, dist_matrix)
avg_silhouette_k1 <- mean(silhouette_result_k1[, 3])  # Calculate the average silhouette width
print(paste("Average Silhouette Score for 5 clusters:", round(avg_silhouette_k1, 4)))
 
# Silhouette analysis for k = 10 clusters
silhouette_result_k2 <- silhouette(k_median_result2$clustering, dist_matrix)
avg_silhouette_k2 <- mean(silhouette_result_k2[, 3])  # Calculate the average silhouette width
print(paste("Average Silhouette Score for 10 clusters:", round(avg_silhouette_k2, 4)))

```

Gaussian Mixture Models (GMM) are probabilistic models that assume data is generated from a mixture of Gaussian distributions. This allows GMM to model complex structures in high-dimensional data, making it more flexible than traditional clustering methods, especially when clusters follow a normal distribution. 
To visualize the clustering results, UMAP is applied for dimensionality reduction before fitting the GMM. UMAP is effective in keeping the local structure of the data while also showing global patterns. This makes it a suitable choice for high-dimensional datasets like the document-term matrix (DTM). Unlike t-SNE, which can distort global relationships, UMAP maintains more meaningful connections among data points and can handle larger datasets efficiently.

```{r}
#| label: GMM Method

set.seed(40)
# Apply UMAP
umap_result <- umap(dtm_matrix)

# GMM with 5 clusters on UMAP
gmm_5_umap <- Mclust(umap_result$layout, G = 5)
summary(gmm_5_umap)

# GMM with 10 clusters on UMAP 
gmm_10_umap <- Mclust(umap_result$layout, G = 10)
summary(gmm_10_umap)

# Convert UMAP results to a data frame
umap_data <- as.data.frame(umap_result$layout)

# Add cluster assignments
umap_data$cluster_5 <- gmm_5_umap$classification
umap_data$cluster_10 <- gmm_10_umap$classification

# Plot for GMM with 5 clusters
ggplot(umap_data, aes(x = V1, y = V2, color = as.factor(cluster_5))) +
  geom_point(alpha = 0.6) +
  labs(title = "GMM Clustering (5 Clusters) - UMAP",
       x = "UMAP Dimension 1",
       y = "UMAP Dimension 2",
       color = "Cluster") +
  theme_minimal()

# Plot for GMM with 10 clusters
ggplot(umap_data, aes(x = V1, y = V2, color = as.factor(cluster_10))) +
  geom_point(alpha = 0.6) +
  labs(title = "GMM Clustering (10 Clusters) - UMAP",
       x = "UMAP Dimension 1",
       y = "UMAP Dimension 2",
       color = "Cluster") +
  theme_minimal()

# Silhouette
# Calculate silhouette scores
silhouette_5_umap <- silhouette(gmm_5_umap$classification, dist(umap_result$layout))
silhouette_10_umap <- silhouette(gmm_10_umap$classification, dist(umap_result$layout))

# Assign the average silhouette scores to variables
avg_silhouette_5_umap <- mean(silhouette_5_umap[, 3])
avg_silhouette_10_umap <- mean(silhouette_10_umap[, 3])

# Print the silhouette scores
cat("Average Silhouette Score for 5 Clusters (UMAP):", avg_silhouette_5_umap, "\n")
cat("Average Silhouette Score for 10 Clusters (UMAP):", avg_silhouette_10_umap, "\n")


# Sentiment labels
sentiment_summary_5 <- as.data.frame(table(umap_data$cluster_5, movie_review$sentiment))
colnames(sentiment_summary_5) <- c("Cluster", "Sentiment", "Count")

sentiment_summary_10 <- as.data.frame(table(umap_data$cluster_10, movie_review$sentiment))
colnames(sentiment_summary_10) <- c("Cluster", "Sentiment", "Count")

print("Sentiment Distribution for 5 Clusters:")
print(sentiment_summary_5)

print("Sentiment Distribution for 10 Clusters:")
print(sentiment_summary_10)

ggplot(sentiment_summary_5, aes(x = factor(Cluster), y = Count, fill = factor(Sentiment))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("red", "green"), labels = c("Negative", "Positive")) +
  labs(title = "Distribution of Sentiments Across Clusters (GMM with UMAP, k = 5)",
       x = "Cluster", 
       y = "Number of Reviews", 
       fill = "Sentiment") +
  theme_minimal()

ggplot(sentiment_summary_10, aes(x = factor(Cluster), y = Count, fill = factor(Sentiment))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("red", "green"), labels = c("Negative", "Positive")) +
  labs(title = "Distribution of Sentiments Across Clusters (GMM with UMAP, k = 10)",
       x = "Cluster", 
       y = "Number of Reviews", 
       fill = "Sentiment") +
  theme_minimal()

# DB Index
db_index_5_umap <- index.DB(umap_result$layout, gmm_5_umap$classification)
db_index_score_5_umap <- db_index_5_umap$DB
print(db_index_score_5_umap)

db_index_10_umap <- index.DB(umap_result$layout, gmm_10_umap$classification)
db_index_score_10_umap <- db_index_10_umap$DB
print(db_index_score_10_umap)

```

Hierarchical Clustering was also applied to our dataset to analyze the natural groupings within the movie reviews. This method calculates the similarity between data points using the Manhattan distance, then organizes them into a hierarchical structure. Initially, each data point is treated as its own cluster. The algorithm iteratively merges the closest clusters based on Manhattan distance, creating a tree-like structure known as a dendrogram. To define specific clusters, we cut the dendrogram at chosen levels, resulting in distinct groups of reviews. This hierarchical approach allows us to capture the nested structure of the data and analyze clusters at multiple levels of granularity.

```{r}

#| label: Hierarchical Clustering


# Step 1: Calculate Distance Matrix
dist_matrix <- dist(dtm_matrix, method = "manhattan") # We could also use "eukleidian"

# Step 2: Perform Hierarchical Clustering
hclust_result <- hclust(dist_matrix, method = "ward.D2") # Using Ward's method for compact clusters

# Set larger plot dimensions (optional, useful for R Markdown)
options(repr.plot.width = 10, repr.plot.height = 6)

# Plot the dendrogram with better formatting
plot(hclust_result, main = "Dendrogram of Movie Review Clusters", 
     xlab = "Reviews", ylab = "Height", cex = 0.6, # Adjust text size
     sub = "", hang = -1) # "hang = -1" makes the labels align at the bottom

# Add rectangles to indicate clusters
rect.hclust(hclust_result, k = 5, border = "red")    # 5 clusters in red
rect.hclust(hclust_result, k = 10, border = "blue")  # 10 clusters in blue



# Step 3: Cut the dendrogram to create 5 and 10 clusters
movie_review$cluster1 <- cutree(hclust_result, k = 5)
movie_review$cluster2 <- cutree(hclust_result, k = 10)

# Step 4: Plot Distribution of Sentiments for 5 clusters
ggplot(movie_review, aes(x = factor(cluster1), fill = factor(sentiment))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("red", "green"), 
                    labels = c("Negative", "Positive")) +
  labs(title = "Distribution of Sentiments Across Clusters (k = 5)",
       x = "Cluster", 
       y = "Number of Reviews", 
       fill = "Sentiment") +
  theme_minimal()

# Step 5: Plot Distribution of Sentiments for 10 clusters
ggplot(movie_review, aes(x = factor(cluster2), fill = factor(sentiment))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("red", "green"), 
                    labels = c("Negative", "Positive")) +
  labs(title = "Distribution of Sentiments Across Clusters (k = 10)",
       x = "Cluster", 
       y = "Number of Reviews", 
       fill = "Sentiment") +
  theme_minimal()

# Step 6: Calculate Silhouette Coefficient
calculate_silhouette <- function(dtm_matrix, clustering_result) {
  silhouette_result <- silhouette(clustering_result, dist_matrix)
  mean(silhouette_result[, 3])  # Average silhouette width
}

# Calculate and print Silhouette Coefficients for 5 and 10 clusters
Hsilhouette_k1 <- calculate_silhouette(dtm_matrix, movie_review$cluster1)
Hsilhouette_k2 <- calculate_silhouette(dtm_matrix, movie_review$cluster2)

# Print Silhouette Coefficients
print(paste("Silhouette Coefficient for 5 clusters:", round(Hsilhouette_k1, 4)))
print(paste("Silhouette Coefficient for 10 clusters:", round(Hsilhouette_k2, 4)))

```

For clustering, Latent Dirichlet Allocation (LDA) was employed to perform topic modeling, with models created for both 5 and 10 topics. LDA is a probabilistic model that assigns each document a mixture of topics, where each topic is defined by a distribution over words. By experimenting with different topic counts (5 and 10), the goal was to find an optimal configuration that best represented the underlying themes in the data.

```{r}

#| label: LDA Topic Modeling

# Step 1: Convert weighted DTM to a matrix for LDA
dtm_for_lda <- as.matrix(dtm)

# Step 2: Fit LDA with 5 topics
lda_5 <- LDA(dtm_for_lda, k = 5, control = list(seed = 123))

# Step 3: Fit LDA with 10 topics
lda_10 <- LDA(dtm_for_lda, k = 10, control = list(seed = 123))

# Step 4: Analyze and Interpret the Results

# Step 5: Get top terms for each topic (for 5 topics)
terms_lda_5 <- terms(lda_5, 10)
print(terms_lda_5)

# Step 6: Get top terms for each topic (for 10 topics)
terms_lda_10 <- terms(lda_10, 10)
print(terms_lda_10)

# Step 7: Assign each document to the most likely topic (for 5 topics)
topics_lda_5 <- topics(lda_5, 1)
table(topics_lda_5)

# Step 8: Assign each document to the most likely topic (for 10 topics)
topics_lda_10 <- topics(lda_10, 1)
table(topics_lda_10)


# Step 9: Assuming you have a sentiment column in movie_review
table(topics_lda_5, movie_review$sentiment)
# Create a data frame with the topics
lda_5_df <- data.frame(document = 1:length(topics_lda_5), topic = topics_lda_5)

# Step 10: Plot the number of documents per topic (for 5 clusters)
ggplot(lda_5_df, aes(x = factor(topic))) +
  geom_bar(fill = "skyblue") +
  labs(title = "Distribution of Documents per Topic (5 Clusters)",
       x = "Topic", y = "Number of Documents") +
  theme_minimal()

# Repeat for 10 clusters
topics_lda_10 <- topics(lda_10, 1)
lda_10_df <- data.frame(document = 1:length(topics_lda_10), topic = topics_lda_10)

ggplot(lda_10_df, aes(x = factor(topic))) +
  geom_bar(fill = "lightcoral") +
  labs(title = "Distribution of Documents per Topic (10 Clusters)",
       x = "Topic", y = "Number of Documents") +
  theme_minimal()

#internal validation

#silhouette 


# Convert topic assignments (clusters) to a factor for silhouette analysis
silhouette_lda_5 <- silhouette(topics_lda_5, dist(dtm_for_lda))

avg_silhouette_lda5 <- mean(silhouette_lda_5[, 3])  # Calculate the average silhouette width
print(paste("Average Silhouette Score for 5 clusters:", round(avg_silhouette_lda5, 4)))

# Repeat for 10 clusters
silhouette_lda_10 <- silhouette(topics_lda_10, dist(dtm_for_lda))

avg_silhouette_lda10 <- mean(silhouette_lda_10[, 3])  # Calculate the average silhouette width
print(paste("Average Silhouette Score for 10 clusters:", round(avg_silhouette_lda10, 4)))


#DaviesBouldin


# Calculate DB-index for LDA with 5 clusters
db_index_5 <- index.DB(dtm_for_lda, topics_lda_5)
db_index_5
table(topics_lda_10, movie_review$sentiment)

# Calculate DB-index for LDA with 10 clusters
db_index_10 <- index.DB(dtm_for_lda, topics_lda_10)
db_index_10

# Access the DB index score directly
db_index_score_5 <- db_index_5$DB
print(db_index_score_5)

db_index_score_10 <- db_index_10$DB
print(db_index_score_10)

```



# Evaluation & model comparison

We use the Silhouette score to compare clustering models because it provides a clear metric to evaluate how well the data points are clustered. The score measures the cohesion within clusters and the separation between different clusters. A higher Silhouette score indicates that data points are well matched to their own clusters and are distinct from other clusters. This makes it an effective way to assess the quality of clustering, especially when different algorithms, such as K-means, K-median, or GMM, are used.

By using the Silhouette score, we can quantitatively determine which model best fits the data. It helps to highlight models where the clusters are more distinct and separated, thus leading to more meaningful patterns. This metric is valuable when trying to choose the optimal clustering approach, as it provides a simple yet effective way to compare models based on their ability to create well-defined clusters.


```{r}
#| label: table example


data.frame(
  model       = c("K-mean", "K-median", "GMM", "Hierarchical Clustering", "LDA Topic Modeling"),
  silhouette = c(avg_silhouette_5,avg_silhouette_k1,avg_silhouette_5_umap,Hsilhouette_k1, avg_silhouette_lda5),
  notes = c("Silhouette Coefficient for 5 clusters")
 
)



data.frame(
  model       = c("K-mean", "K-median", "GMM", "Hierarchical Clustering","LDA Topic Modeling"),
  silhouette = c(avg_silhouette_10,avg_silhouette_k2, avg_silhouette_10_umap, Hsilhouette_k2, avg_silhouette_lda10 ),
  notes = c("Silhouette Coefficient for 10 clusters")
 
)
```



